Cost- and Deadline-Constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-04-25 paper/paper.tex
intro and related work

Scientific workflows, usually represented as directed acyclic graphs (DAGs), are
an important class of distributed applications that have been studied in the context of
resource management and task scheduling on grid and utility computing systems.
Workflow applications for large computational problems are often composed of 
several inter-related workflows grouped into {\em ensembles}. All the
workflows in an ensemble typically have a similar structure, but they differ in their
input data, number of tasks, and individual task sizes. The CyberShake 
application~\cite{Callaghan2011} is a good example of an application that requires 
scientific workflow ensembles. CyberShake computes seismic hazard curves for a given
geographical location. In order to produce a hazard map for a region such as 
Southern California, many hazard curves need to be computed for a large number of 
geographical sites, where each site requires running a single, complex scientific 
workflow. Workflows for individual sites may differ not only in their parameters, 
but also in their priority. For example, some sites may be in heavily populated areas 
or in strategic locations such as power plants, while others may be less important. 
Other examples of workflow ensembles come from astronomical applications that create
science-grade image mosaics~\cite{Deelman2008} or search for extrasolar 
planets~\cite{Vockler2011}, where multiple, different workflows are used to
process data for different areas of the sky.

@@

In this paper, we aim to gain insight into resource management challenges when executing 
scientific workflow ensembles on clouds. In particular, we address a new and important 
problem of maximizing the number of completed workflows from an ensemble under both budget and
deadline constraints. The motivation for this work is to answer the fundamental question of 
concern to a researcher: How much computation can be completed given the limited budget and 
timeframe of a research project? The goals of this paper are to discuss and assess possible
static and dynamic strategies for both task scheduling and resource
provisioning. Based on the knowledge of workflow scheduling algorithms we
analyze strategies for on-line scheduling of individual tasks as well as static
algorithms that rely on the information about the workflow structure (critical
paths and workflow levels) and estimates of task runtimes. In addition, we
analyze a hybrid workflow-aware dynamic scheduling algorithm, which uses the
workflow structure information to estimate which workflows should be rejected
from the ensemble based on the constraints. We evaluate these algorithms using a
simulator we have developed using CloudSim~\cite{Calheiros2011}, which models the 
infrastructure and the application. The simulator includes also tightly-coupled 
scheduler and provisioner modules. The algorithms were evaluated on a set 
of scientific workflow ensembles from several different applications using a 
broad range of budget and deadline parameters.

The main contributions of this paper are:
\begin{itemize}
  \item a definition of the problem of scheduling prioritized workflows grouped into
  ensembles, under both budget and deadline constraints,
  \item an analysis of static and dynamic strategies for task scheduling and
  resource provisioning on clouds, together with admission procedures to
  predict which workflows from the ensemble can be completed,
  \item the development of several static and dynamic algorithms and of a simulator based
  on the defined application and infrastructure model,
  \item a performance evaluation of the algorithms on a set of synthetic benchmark
  workflow ensembles based on important, real scientific applications,
  \item an analysis of performance results that assesses the advantages and
  disadvantages of the proposed algorithms on different application scenarios and
  varying constraint values.
\end{itemize}

The paper is organized as follows: after an analysis of related work in
Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
infrastructure and application model we are targeting in this paper.
In Section~\ref{sec:algorithms} we describe the dynamic and
static algorithms we developed. Section~\ref{sec:performance} presents test
scenarios, performance metrics, and the results of simulation studies of these algorithms.
Finally, general conclusions, lessons learned, and future work are outlined in
Section~\ref{sec:conclusions}.

\section{Related Work}
\label{sec:related}
General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
Auto Scaling\footnote{\url{http://aws.amazon.com/autoscaling}} and
RightScale\footnote{\url{http://www.rightscale.com}}) allow adjusting the size
of resource pool based on metrics related to infrastructure and application.
A typical infrastruc\-ture-specific metric is system load, whereas
application-speci\-fic metrics include response time and length of a task or
of a request queue. It is possible to set thresholds and limits to tune the behavior
of these autoscaling systems, but no support for complex applications is provided.

@@

\section{Discussion of Results}

>>>>

Scientific workflows, usually represented as directed acyclic graphs (DAGs), are
an important class of distributed applications that lead to challenging problems
in resource management on grid and utility computing systems. Workflow
applications for large computational problems are often composed of several
inter-related workflows grouped into {\em ensembles}. All the workflows in an
ensemble typically have a similar structure, but they differ in their input
data, number of tasks, and individual task sizes.

The CyberShake~\cite{Callaghan2011} is a good example of an application that
requires scientific workflow ensembles. Generating a hazard map for a region
such as Southern California requires an ensemble of many related workflows to
compute individual hazard curves. The hazard map generated by the CyberShake
group in 2009, for example, required an ensemble of 239 workflows. In astronomy,
for example, a user might wish to use multiple Montage~\cite{Deelman2008}
workflows with different parameters to generate a set of image mosaics that
cover a given area of the sky in different wavelengths, or use an ensemble to generate
the parts of a very large mosaic. The Pegasus team is now working on a Galactic
Plane project that is focused on generating a mosaic of the entire sky. This
ensemble consists of 17 workflows, each of which contains 900 sub-workflows and
computes one ``tile'' of the full mosaic. Another ensemble example is the
Periodograms application~\cite{Vockler2011} that searches for extrasolar planet
by detecting the periodic dips in light intensity. Due to the large scale of the
input data, this application is often split up into multiple batches processed
by different workflow. Additional workflows are created to run the analysis
using different parameters. A recent analysis of data from the Kepler satellite
required three ensembles of 15 workflows.

Ensemble workflows may differ not only in their parameters, but also in their
priority. For example, in CyberShake some sites may be in heavily populated
areas or in strategic locations such as power plants, while others may be less
important. Scientists typically prioritize the workflows in an ensemble such
that important workflows are finished first. This enables them to see critical
results early, and helps them to choose the most important workflows when the
time and financial resources available for running the computations are limited.

@@

In this paper, we aim to gain insight into resource management challenges when
executing scientific workflow ensembles on clouds. In particular, we address a
new and important problem of maximizing the number of completed workflows from
an ensemble under both budget and deadline constraints. The motivation for this
work is to answer the fundamental question of concern to a researcher: How much
computation can be completed given the limited budget and timeframe of a
research project?

The main contributions of this paper are:
\begin{itemize}
  \item we define the problem of scheduling prioritized workflows grouped
  into ensembles, under both budget and deadline constraints,
  \item we analyze and develop dynamic (on-line) strategies for task scheduling
  and resource provisioning as well as static algorithms that
  rely on the information about the workflow structure (critical paths and
  workflow levels) and estimates of task runtimes. 
  \item we analyze and develop a hybrid workflow-aware dynamic scheduling
  algorithm, together with admission procedures to predict which workflows from
  the ensemble can be completed,
  \item we evaluate these algorithms using a simulator we have
  developed using CloudSim~\cite{Calheiros2011}, which models the infrastructure
  and the application, taking into account uncertainties in task runtime
  estimations, resource provisioning delays and failures, 
  \item we discuss the performance of the algorithms on a set of
  synthetic benchmark workflow ensembles based on important, real scientific
  applications, in the presence of such of such factors as 
  uncertainties in task runtime estimations, resource provisioning delays and
  failures, on a broad range of different application scenarios and
  varying constraint values.
\end{itemize}

The paper is organized as follows: after an analysis of related work in
Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
infrastructure and application model we are targeting in this paper. In
Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
developed. Section~\ref{sec:performance} presents test scenarios and performance
metrics, while Section~\ref{sec:results} discusses the results of simulation
studies of these algorithms. Finally, general conclusions, lessons learned, and
future work are outlined in Section~\ref{sec:conclusions}.

\section{Related Work}
\label{sec:related}
General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
Auto Scaling~\cite{Autoscaling} and
RightScale~\cite{RightScale}) allow adjusting the size
of resource pool based on metrics related to infrastructure and application.
A typical infrastruc\-ture-specific metric is system load, whereas
application-speci\-fic metrics include response time and length of a task or
of a request queue. It is possible to set thresholds and limits to tune the behavior
of these autoscaling systems, but no support for complex applications is provided.

@@

\section{Discussion of Results}
\label{sec:results}
